---
title: "analyzeRiksdag-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analyzeRiksdag-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(analyzeRiksdag)
```

# Riksdag API
## Description
> Sveriges Riksdag is the highest decision-making assembly in Sweden. Possible topics are categorized into parliamentary committees (utskott), such as Education, Defence, Health and Welfare and many more. When these committees vote, topics are divided into designations that are identified with the committee abbreviation and a number (i.e. AU10), and within these designations several agenda points may be voted on. The Riksdag provides an API that allows the user to access data on how party members voted on each of these points as far back as 2002. This package provides some functions that allow the user to interact with this API and do some simple analysis of the data.

## Functionality
> The API can be queried through a form that asks the user the fill out their choices for several variables. These are then appended to the URL, which, depending on the chosen data format, will download or load the data in browser. In this package, only the assembly year, designation and agenda point may be specified and data will be downloaded for all parties, regions, members, and types of answers. This allows the user to further filter the data as desired after fetching, while not being overwhelmed by the required input. 

# Functions and examples
> To fetch data from the Riksdag API, `fun_fetch_data()` may be used. 
